---
title: "Prediction Assignment Writeup"
output:
  html_document:
    df_print: paged
---
###Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

###Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

##Introduction and Loadind dataset

```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(gbm)
library(plyr)
dt_training <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
dt_testing <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```

##Exploratory analysis and cleaning data
```{r}
str(dt_testing)
head(dt_testing)
str(dt_training)
head(dt_training)
```

In this section, we will remove all columns that contains NA and remove features that are not in the testing dataset. The features containing NA are the variance, mean and standard devition (SD) within each window for each feature. Since the testing dataset has no time-dependence, these values are useless and can be disregarded. We will also remove the first 7 features since they are related to the time-series or are not numeric.
```{r}
features <- names(dt_testing[,colSums(is.na(dt_testing)) == 0])[8:59]

# Only use features used in testing cases.
dt_training <- dt_training[,c(features,"classe")]
dt_testing <- dt_testing[,c(features,"problem_id")]

dim(dt_training); dim(dt_testing);
```

##Partitioning the Dataset
As recomended in course: we will split our data into a training data set (60% of the total cases) and a testing data set (40% of the total cases; the latter should not be confused with the data in the pml-testing.csv file). This will allow us to estimate the out of sample error of our predictor.
```{r}
set.seed(1234)

inTrain <- createDataPartition(dt_training$classe, p=0.6, list=FALSE)
training <- dt_training[inTrain,]
testing <- dt_training[-inTrain,]

dim(training); dim(testing);
```

##Building the Decision Tree Model
```{r}
set.seed(1234)
modFitDT <- rpart(classe ~ ., data = training, method="class", control = rpart.control(method = "cv", number = 10))
fancyRpartPlot(modFitDT ,sub = ("Decision Tree Model 14-Jul-2020"))
```

###Predicting with the Decision Tree Model
```{r}
set.seed(1234)

prediction <- predict(modFitDT, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```

##Building the Random Forest Model
```{r}
set.seed(1234)

modFitRF <- randomForest(classe ~ ., data = training, method = "rf", importance = T, trControl = trainControl(method = "cv", classProbs=TRUE,savePredictions=TRUE,allowParallel=TRUE, number = 10))

plot(modFitRF)
```

##Building the Boosting Model
```{r}
set.seed(1234)
modFitBoost <- train(classe ~ ., method = "gbm", data = training,
                    verbose = F,
                    trControl = trainControl(method = "cv", number = 10))

modFitBoost
plot(modFitBoost)
```

##Predicting with the Random Forest Model
```{r}
prediction <- predict(modFitRF, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```

##Predicting with the Boosting Model
```{r}
prediction <- predict(modFitBoost, testing)
confusionMatrix(prediction, testing$classe)
```

#Predicting with the Testing Data (pml-testing.csv)
##Decision Tree Prediction
```{r}
predictionDT <- predict(modFitDT, dt_testing)
predictionDT
```
##Random Forest Prediction
```{r}
predictionRF <- predict(modFitRF, dt_testing)
predictionRF
```
##Boosting Prediction
```{r}
predictionBoost <- predict(modFitBoost, dt_testing)
predictionBoost
```

